Linear response theory for core set identification

### 2012-07-22 writedown/linear.response/langevin.tex
new figs

\subsection{Numerical test}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-pot.eps}
  \caption{The potential with and without perturbation.}
  \label{fig:tmp1}
\end{figure}

We test the idea by a one-dimensional model system: one particle in a
double-well potential. For convenience, we let the mass of the
particle to be 1 \textsf{amu}. The Hamiltonian of the system is given
by:
\begin{align}
  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
\end{align}
with
\begin{align}
  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
\end{align}
Here $k = 8$ $\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1 \textsf{nm}$.
Notice at room temperature $300 \textsf{K}$, $k_BT = 2.48$ \textsf{kJ/mol}.
See the red line in Fig.~\ref{fig:tmp1}.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is given by
\begin{align}
  \vect C(\vect p, \vect q) = 0; \qquad
  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
\end{align}
Here $V(\vect q) = -\vect q$, which effectively tilt the original
potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
by the following function:
\begin{align}
  F_e(t) = 
  \begin{cases}
    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
    F_e^{\textrm{max}} & t \geq t_c
  \end{cases}
\end{align}
$t_c = 100$ \textsf{ps} so that the perturbation increases slow
enough: the typical decaying time scale of the correlation function
in~\eqref{eqn:main} is roughly 3 \textsf{ps}.  We choose $\mt_\alpha$
to be the propagator of the Langevin dynamics at temperature $150$
\textsf{K} with $\alpha = 1$ \textsf{ps}.  Since the dynamics is in
1-d, the projection is identity.  We consider the stability of the
distribution $f_t$ at time $t$: $-\Delta\mt_\alpha f_t$. The smaller
this value, the more stable the core sets are.

See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
the direct non-equilibrium simulation and the linear response
approximation are presented. Time slices $t = 0$, 30, 60 and 100
\textsf{ps} are shown.  When the perturbation is small,
i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
and the linear response results consistent very well. As the
perturbation grows stronger, the left well becomes less stable, while
the right well becomes more stable.  When the perturbation is big,
i.e. $F_e^{\textrm{max}} = 6$, the linear response results are no
longer precise. When $t\geq 60$ \textsf{ps}, the left well basically
disappears, however, the linear response calculation present a well
that is nearly as stable as the right well.

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/warm100.str2.0.nst1e09.smallgrid/fig-2d.eps}
  \caption{The stability of distribution $f_t$ under perturbation of
    $F_e^{\textrm{max}} = 2$.
    Upper row: the direct non-equilibrium simulation. Lower row: the
    results of linear response formula~\eqref{eqn:main}.
  }
  \label{fig:tmp2}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/warm100.str6.0.nst1e09.smallgrid/fig-2d.eps}
  \caption{The stability of distribution $f_t$ under perturbation of
    $F_e^{\textrm{max}} = 6$.  Upper row: the direct non-equilibrium
    simulation. Lower row: the results of linear response
    formula~\eqref{eqn:main}.  }
  \label{fig:tmp3}
\end{figure}

>>>>

\subsection{Numerical tests}

\subsubsection{The tilting double-well potential}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
  \caption{The double-well potential with tilting perturbation.}
  \label{fig:tmp1}
\end{figure}

We test the idea firstly by a one-dimensional model system: one particle in a
tilting double-well potential. For convenience, we let the mass of the
particle to be 1 \textsf{amu}. The unperturbed
Hamiltonian of the system is given by:
\begin{align}
  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
\end{align}
with potential
\begin{align}
  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
See the red line in Fig.~\ref{fig:tmp1}.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is given by
\begin{align}
  \vect C(\vect p, \vect q) = 0; \qquad
  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
\end{align}
Here $V(\vect q) = -\vect q$ is  effectively tilt the original
potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
by the following function:
\begin{align}
  F_e(t) = 
  \begin{cases}
    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
    F_e^{\textrm{max}} & t \geq t_c
  \end{cases}
\end{align}
$t_c = 20$~\textsf{ps} so that the perturbation increases slow
enough: the typical decaying time scale of the correlation function
in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
We choose $\tau = 1$~\textsf{ps}.
% We choose $\mt_\alpha$
% to be the propagator of the Langevin dynamics at temperature $150$
% \textsf{K} with $\alpha = 1$ \textsf{ps}.
Since the dynamics is in
1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
The larger
this value, the more stable the core sets are.

See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
the direct non-equilibrium simulation and the linear response
approximation are presented. Time slices $t = 0$, 10, 15 and 20
\textsf{ps} are shown.  When the perturbation is small,
i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
and the linear response results are accurately consistent. As the
perturbation grows stronger, the left well becomes less stable, while
the right well becomes more stable.  When the perturbation is big,
i.e. $F_e^{\textrm{max}} = 4$, the linear response results are no
longer precise. When $t\geq 10$ \textsf{ps}, the left well basically
disappears, however, the linear response calculation presents
artifical occilation at the right well.

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/tilt.warm020.str2.0.nst1e09.smallgrid/fig-2d.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 2$.
    Upper row: the direct non-equilibrium simulation. Lower row: the
    results of linear response formula~\eqref{eqn:core-identify-approx-2}.
  }
  \label{fig:tmp2}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/tilt.warm020.str4.0.nst1e09.smallgrid/fig-2d.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 4$.  Upper row: the direct non-equilibrium
    simulation. Lower row: the results of linear response
    formula~\eqref{eqn:core-identify-approx-2}.  }
  \label{fig:tmp3}
\end{figure}

\subsubsection{Splitting single-well potential}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
  \caption{The single-well potential with splitting perturbation.}
  \label{fig:tmp4}
\end{figure}

In this splitting single-well potential example, the unperturbed
potential is simply the harmonic potential given by:
\begin{align}
  U(\vect q) = \frac12\,k\,\vect q^2 
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
See the red line in Fig.~\ref{fig:tmp4}.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is again given by
\begin{align}
  \vect C(\vect p, \vect q) = 0; \qquad
  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
\end{align}
where $V(\vect q)$ is a Gaussian type potential:
\begin{align}
  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
\end{align}
we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/split.warm020.str1.0.nst1e09.smallgrid.more/fig-2d.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 1.0$.
    Upper row: the direct non-equilibrium simulation. Lower row: the
    results of linear response formula~\eqref{eqn:core-identify-approx-2}.
  }
  \label{fig:tmp5}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/split.warm020.str2.0.nst1e09.smallgrid.more/fig-2d.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 2.0$.  Upper row: the direct non-equilibrium
    simulation. Lower row: the results of linear response
    formula~\eqref{eqn:core-identify-approx-2}.  }
  \label{fig:tmp6}
\end{figure}
